---
title: Assign titles in plotting using mapply()
author: Yukai Zou
date: '2016-04-04'
slug: assign-titles-in-plotting-using-mapply
categories:
  - R
tags:
  - base
  - plot
---

I often encounter the need of creating figures with subplots, where manually assign titles can become a time-consuming task. In R base plotting, I can quickly assign titles (plus other asthestics) by using the `mapply()` function:

- `mapply(function, target, parameters)`

Here I use the `iris` dataset as an example:

```{r, echo=FALSE}
head(iris)
```

I'd like to plot out histograms to look at the distributions, and at this stage, I don't care much about the last column (`Species`).

```{r}
# Initialize the plot with 2 rows and 2 columns for subplotting
par(mfrow=c(2,2))

# Plot out histograms for the first 4 columns of iris
mapply(hist, iris[,1:4], 
       # Get the column names and assign as x label
       xlab=colnames(iris[,1:4]), 
       # Assign "Histogram of ..." as title name (main)
       main=paste0("Histogram of ", colnames(iris[,1:4])))
```

One `mapply()`, save me tons.
