---
title: Math notation and customised ticks in ggplot2
author: Yukai Zou
date: '2016-12-07'
slug: math-expression-and-customised-ticks-in-ggplot2
categories:
  - R
tags:
  - ggplot2
---

```{r setup, include=F}
library(ggplot2); library(dplyr)
```


# Adding math notation

On some occasions, one might need to add math notations on a plot. To do so, add `expression()` to the ggplot expression.

Consider this very simple example:

```{r, echo=F}
dat = data.frame(x = c(1, 2, 3), 
                 y = c(10, 20, 30), 
                 y2 = c(10^2, 20^2, 30^2))
dat
```
What `y2` means here is essentially "y square". Plotting x and y is relatively straightforward:

```{r}
ggplot(dat, aes(x = x, y = y)) + geom_point()
```


But in order to plot `y2`, one would consider updating the name of y-axis to avoid confusion, in this case, making 2 a superscript to become $y^{2}$.

```{r}
ggplot(dat, aes(x = x, y = y2)) + geom_point() +
        ylab(expression(y^{2}))
```

Note that the math expression cannot be bolded.

# Defining customized ticks

This can be achieved by adding the following functions to the ggplot expression, depending on which axis (x/y) and the type of variable (discrete/continuous):

- `scale_x_discrete()`
- `scale_y_discrete()`
- `scale_x_continuous()`
- `scale_y_continuous()`

where the `break` attribute would need to be specified.

Let's work on one of the examples above that plots out $y^2$. I'd like to specify the y ticks so that it shows 900 on top.

```{r}
ggplot(dat, aes(x = x, y = y2)) + geom_point() +
        ylab(expression(y^{2})) +
        scale_y_continuous(breaks = c(300, 600, 900))

```

