---
title: About spread() in tidyr
author: Yukai Zou
date: '2016-09-13'
slug: about-spread-in-tidyr
categories:
  - R
tags:
  - tidyr
---

```{r setup, include=F}
library(dplyr); library(tidyr)
```
**Update**: now we have `pivot_wider()` which works effectively the same. Check it out [here](https://tidyr.tidyverse.org/reference/pivot_wider.html).

`spread()` can reshape dataset containing measures of multiple time points and help query data availability per subject, which can truly be a time saver when there are many subjects with lots of data. 

- `spread(data, key, value)`

Now, consider the following (minimal) dataset that has 3 subjects and measures from 3 time points: 

```{r, echo=F}
dat <- data.frame(Subject = c(rep("A", 3), rep("B", 3), rep("C", 3)),
                  Timepoint = rep(c("T1", "T2", "T3"), 3),
                  Measure = c(10, 12, NA, 9, 10, 9, 6, NA, NA))
dat
```
Unless reading line by line, it's not immediate to tell whether the measure at a certain timepoint is missing or not. It gets even more cumbersome when the list is long. What if say we transform the table to let the `Timepoint` becomes separate columns, and the values within `Measure` are stored within those columns instead? This is where `spread()` can be applied:

```{r}
dat %>% spread(key = Timepoint, value = Measure)
```
It gets easier to read along each row to grasp the idea of data availability, plus the dataset [becomes much more tidy](https://vita.had.co.nz/papers/tidy-data.pdf).
