---
title: Convert p-values using stars.pvals() and cut()
author: Yukai Zou
date: '2016-03-25'
slug: convert-p-values-using-stars-pvals-and-cut
categories:
  - R
tags:
  - gtools
  - base
---

```{r setup, include=FALSE}
library(gtools)
```

Consider the scenario where we run some statistical analyses using R. We get a bunch of p-values, and we need to highlight those that are significant, so they don't all look the same. There are two functions that could help automate this tedious task.

# stars.pval()

`stars.pval()` from `{gtools}` package can convert p-values to stars.

Let's simulate some p-values:

```{r}
x <- seq(from = 0.001, to = 0.1, by = 0.005)
x
```

and apply `stars.pval()`:

```{r}
stars.pval(x)
```

The output also returns the legends, which is a nice feature that explains what p-value ranges are mapped to what symbols.

# cut()

The `cut()` function can divide numbers into several sections, which could act as an extension to `star.pvals()` who just converts p-values to stars. 

- By specifying the `break` attribute, the numeric vector can be divided either evenly or according to what's defined in the vector.

Working example:

Consider the same simulated vector above:

```{r, echo=FALSE}
x
```

To break this vector evenly into 10 subsets (bins):

```{r}
cut(x, breaks = 10)
```
Note that the output is a _factor_ datatype that comes with 10 levels.

To cut the vectors according to p-value ranges:

```{r}
cut(x, breaks = c(0, 0.001, 0.01, 0.05, 1))
```
Putting all the outputs to a data frame:

```{r}
data.frame(pval = x,
           cut.pval = cut(x, breaks = c(0, 0.001, 0.01, 0.05, 1)),
           stars.pval = stars.pval(x))
```
