---
title: Notes on string editing
author: Yukai Zou
date: '2017-06-17'
slug: notes-on-string-editing
categories:
  - R
tags:
  - string
---

# nchar(): get rid of the last few digits

Sometimes one may wish to extract substrings in a character vector from the end rather than the beginning, and a workaround is to use the `nchar()` function.

Working example:

```{r}
text <- "12345678"
text

nchar(text)
```

To keep the last four characters:

```{r}
substr(text, nchar(text) - 3, nchar(text))
```

<a name="add-boundary"></a>
# Adding boundaries - "\b", "^", "$"

These operators help identify the exact matched pattern, which is helpful when used in combination of `grep()` and `gsub()` functions. 

For instance:

```{r}
string1 <- c("java", "javascript")

grep("java", string1)
```

As you can see, without specifying boundary, the output indicates that both `java` and `javascript` are returned. We'll fix that by using `\b`.

```{r}
grep("\\bjava\\b", string1)
```

So now we only get `java` returned as the output. Also, note that an extra `\` is needed in the expression.

But there's one thing that `\b` does not recognize: _space_. See this example:

```{r}
string = c("yourtext", "yourtexts", "yourtext ", " yourtext ")

grep("\\byourtext\\b", string)
```

The output tells us that `\\byourtext\\b` has found not only the 1st entry ("`yourtext`"), but also the 3rd ("`yourtext `") and 4th (<code> yourtext </code>") entries as well - didn't work as expected. 

To fix that problem, `^` and `$` can be used instead. They acted like a _forced_ boundary for the start (`^`) and end (`$`) point.

```{r}
string = c("yourtext", "yourtexts", "yourtext ", " yourtext ")

grep("^yourtext$", string)
```

**See also:** [Extract characters from string](/2016/02/20/extract-characters-from-string/), [This post](https://stackoverflow.com/questions/26813667/how-to-use-grep-gsub-to-find-exact-match) on Stack Overflow.
