---
title: melt() to plot via ggplot2
author: Yukai Zou
date: '2016-04-05'
slug: melt-to-plot-via-ggplot2
categories:
  - R
tags:
  - ggplot2
  - reshape2
  - dplyr
---

```{r setup, include=FALSE}
library(ggplot2); library(reshape2); library(dplyr)
```


# A good table can be visualised quickly

The more I work on data visualization using R, the more I realized that a _good_ table has a clean structure and requires much less effort to be plotted out.

In `ggplot`, this often means that if a table requires more than a `melt()` in `{reshape2}` package then the table is not good enough - needs some more cleaning.

Working example:

```{r}
d1 <- iris
d1 %>% head()
```

```{r}
d1_melted <- d1 %>% melt(id.vars = c("Species"))
d1_melted %>% head()
```

The `facet_wrap()` function can create subplots according to defined attributes. Here, I'd like to:

 - display different `Species` vertically, and 
 - along each row to plot out length and with of sepal and petal. Now that these data are stored in `variable` and `value` from the `melt()`, they can be specified in ggplot expression.

```{r}
d1_melted %>% ggplot(aes(x=value)) + geom_histogram() +
        facet_wrap(Species ~ variable) +
        theme_bw()
```

**Take home message:** Be goal-oriented and thinking backward. Better-engineered data allows visualization and decision-makingm more efficiently.
