---
title: Collapse multiple elements in a vector using paste()
author: Yukai Zou
date: '2016-04-14'
slug: collapse-multiple-elements-in-a-vector-using-paste
categories:
  - R
tags:
  - base
---

To do so, specify the `collapse` attribute in `paste()` function:

- `paste(..., collapse = "")`

Collapsing elements can be quite useful, not only for text processing and cleaning purposes, but also filtering a dataset to narrow it down to observations that are of interest. Of course, there are a ton of tools out there to achieve this (e.g. `dplyr::filter`) but to me it's more natural sometimes to just _write out_ what I want to look at in a string vector first. I'll explain with an example.

Consider the classic `mtcars` dataset:

```{r, echo=F}
head(mtcars, n = 10)
```
There are many models here, and in fact I only need to look at `Mazda RX4` and `Valiant`. 

```{r}
# Code "Mazda RX4" and "Valiant" in a string vector
string_model <- c("^Mazda RX4$", "Valiant")

# Collapse the two elements into one
string_model_collapse <- paste(string_model, collapse = "|")
string_model_collapse
```

Note that `^` and `$` were used to define boundary at the start and the end, so that only "Mazda RX4" will be grabbed and not "Mazda RX4 Wag". See [this blog](/2017/06/17/notes-on-string-editing/index.html#add-boundary) for details. 

Now I have a single-element string vector, Next is to plug the new vector into the `grep()` function:

```{r}
id_model_of_interest <- grep(string_model_collapse, row.names(mtcars))
id_model_of_interest
```
Note that the model names are not stored in any column, which needs to be retrived using `row.names()` function.

```{r}
mtcars[id_model_of_interest,]
```

