---
title: Examine vectors using %in% operator
author: Yukai Zou
date: '2016-04-05'
slug: examine-vectors-using-in-operator
categories:
  - R
tags:
  - base
---

The `%in%` operator allows comparing two vectors by identifying whether the element of one vector exists in the other.

The rule here is that the vector _to be examined_ will come first, followed by `%in%` and then the target vector _to be compared with_.

Working example:

```{r}
vector1 <- c("Individual 1", "Individual 2", "Individual 3")
vector2 <- c("Individual 1", "Individual 2")

# Examine whether elements of vector1 exist in vector 2
vector1 %in% vector2
```

The outputs are logical values with respect to each element in `vector1`. Such output can be useful in vector subsetting:

```{r}
# Get the elements in vector 1 that exist in vector 2
vector1[ vector1 %in% vector2 ]
```


To find out the _opposite_ (i.e. which element is absent), wrap the whole expression with an exclamation and a set of parentheses:

<center><code>!( expression )<code></center>

```{r}
!(vector1 %in% vector2)
```
