---
title: '"stringAsFactors = FALSE": Let strings be strings'
author: Yukai Zou
date: '2016-03-29'
slug: stringasfactors-false-let-strings-be-strings
categories:
  - R
tags:
  - base
---

**Update:** Since `R 4.0.0`, the default for `stringAsFactors` has been set to `FALSE`. It's a good idea to be explicit, so that the code can run more robustly across different R versions.

One of the default behaviors in older versions of R, which is sometimes too aggressive, is converting strings to a datatype called `factor`.

```{r}
t1 <- data.frame(label = rep(c("A", "B"), 5), stringsAsFactors = TRUE)
t1$label
```

If a new string value is entered:

```{r}
t1$label[1] <- "C"
```

R won't recognize the string and instead generated a missing value. To resolve this issue: 

```{r}
t2 <- data.frame(label = rep(c("A", "B"), 5), stringsAsFactors = FALSE)
t2

t2$label[1] <- "C"
t2
```
The column can be transformed into `factor` later if necessary:

```{r}
t2 <- transform(t2, label2 = factor(label))
t2$label2
```

**See also:** [R-blogger post](https://www.r-bloggers.com/2018/03/r-tip-use-stringsasfactors-false/)
